---
title: "Practical Machine Learning Course Project"
author: "Gustavo Ogasawara"
date: "January 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

Human Activity Recognition (HAR) data were used to develop a prediction model that predicts the class of a movement based on a series of measured parameters.

After cleaning process, the data was divided in train and test sets, and then random forest method was applied to construct a prediction model for 5 classes (A, B, C, D, E).

The model predicts the test set with 99% of accuracy and OOB estimated of error rate was 0.84%.

Thanks for Groupware to provide the data: http://groupware.les.inf.puc-rio.br/har#sbia_paper_section


## Downloading Data

```{r download}
if (!file.exists("pml-training.csv")){
        download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile =     "pml-training.csv", method = "wget")}
if (!file.exists("pml-testing.csv")){
        download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile =     "pml-testing.csv", method = "wget")
}

datatrain <- read.csv("pml-training.csv", na.strings = c("NA","#DIV/0!",""))
testquiz <- read.csv("pml-testing.csv", na.strings = c("NA","#DIV/0!",""))
```

## Cleaning Data

To remove NA columns and non-predictors as index, subjects.. etc.

```{r cleaning}
#removing NA columns
datatrain <- datatrain[,colSums(is.na(datatrain)) == 0]
testquiz <- testquiz[,colSums(is.na(testquiz)) == 0]

#removing non-predictors (index, subjects..)
datatrain<-datatrain[,-c(1:7)]
testquiz<-testquiz[,-c(1:7)]

dim(datatrain)
dim(testquiz)

```

## Partition Training Data

Creating train and test sets.

```{r partitioning}

library(caret)

set.seed(666)

inTrain <- createDataPartition(datatrain$classe, p = 0.6, list = FALSE)

train   <- datatrain[inTrain,]
test    <- datatrain[-inTrain,]


```

## Prediction Model

Using random forest method with Parallel to reduce computation time as seeing in Class Forum.

```{r modeling}

library(randomForest)
library(e1071)
library(parallel)
library(doParallel)

modname <- "mymodfit.RData"

if (!file.exists(modname)) {

    cores <- makeCluster(detectCores() - 1)
    registerDoParallel(cores)
    getDoParWorkers()     
    
    
    modfit <- train(classe ~ ., data = train, method = "rf", metric = "Accuracy", trControl=trainControl(method = "cv", number = 4, p= 0.60, allowParallel = TRUE))

    save(modfit, file = "mymodfit.RData")
   
    stopCluster(cores)
    
} else {
    
    load(file = modname, verbose = TRUE)
}


```

## Accuracy and validation tests



```{r}

predtest <- predict(modfit, test)
confusionMatrix(predtest, test$classe)

```
The model worked with high accuracy and low error.

```{r}
modfit$finalModel
```
## Checking Most Important Variables

```{r}
varImp(modfit)
```

## Predicts for Quiz
```{r}
predtestquiz <- predict(modfit, testquiz)
predtestquiz
```


